\pkg{cleanR}: Maid for Cleaning Datasets in \proglang{R}

### 2016-12-15 latex/article_vol2.tex
More structure, finished examples.

@@

By default, a pdf overview document is produced, saved to the disc and opened for inspection. Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts. First, an overview of what was done is presented under the title \textit{Data cleaning summary}. Secondly, each variable in the dataset is presented in turn using (up to) three tools in the \textit{Variable list}: A table summarizing key features of the variable, a figure visualizing its distribution and potentially also a list of flagged issues. For instance, in the \R{numeric}-type variable \R{var2} from \R{toyData}, \R{clean()} has identified two values that are suspected to be miscoded missing values (\R{999} and \R{NaN}), while two values were also flagged as potential outliers that should be investigated more carefully. We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable. \hl{hm, delete the last part? or use different variable? Only these exact two checks were performed}.

Of course, 
\\

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=8cm,page=2]{cleanR_toyData.pdf} 
\includegraphics[width=8cm, page=3]{cleanR_toyData.pdf} 
%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}

\end{center}
\label{fig:example1}
\caption{Example output from running XXXX on the testData
  dataset. First a summary of the full dataset,.. 
  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically we want the .html-version of this file printed as pdf.}}
\end{figure}

\subsection{Controlling contents}
\begin{table}
\small
\begin{tabular}{p{0.2\linewidth}p{0.5\linewidth}p{0.2\linewidth}}
\hline
Argument & Description & Default value \\
\hline

@@

\hline
\end{tabular}
\caption{A selection of commonly used arguments of \R{clean()}.  \hl{I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions}}
\label{table.cleanFormals}
\end{table}

By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset. There are three stages in which this occurs \hl{(with ref. to figure/flowchart?)}:
\begin{enumerate}
\item In the precheck functions
\item In the summarize/visualize/check (SVC) step
\item In the multivariate visuzalizations \hl{Or whatever, in the stuff that we have not yet implemented. I'm pretending like this stage doesn't exist below.}
\end{enumerate}
Each of these stages are controllable using appropriate function parameters in \code{clean}. In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs. However, if for instance the dataset at hand requires completely different visualizations, more control is needed. \pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s. They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes. By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands. Therefore, this section is dedicated to introducing each in turn.\\

\hl{Something like a figure that gives an overview of all summaryFunctions, ... available in standard cleanR, including default settings and some kind of description of where they are called (in precheck or SVC step). Also something about what datatypes they can each be called on. Also, more text here about the relationship between the two (three) stages and the three function types.}

@@

\begin{table}[h]
\small
\centering
\begin{tabular}{p{0.3\linewidth} p{0.35\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
 p{0.01\linewidth} p{0.01\linewidth}}
  \hline
& Description &  \multicolumn{7}{c}{Relevant variable classes} \\ \smallskip
 & &  C & F & I & L & B & N & D\\ 
  \hline \smallskip
  \textbf{\R{summaryFunction}s}  \smallskip \\
  \quad \R{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\ 
  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ &  \\ 
  \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \smallskip \\
 \textbf{\R{visualFunction}s} \smallskip \\
  \quad \R{basicVisual} & Histograms and barplots using graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \smallskip \\
 \textbf{\R{checkFunction}s} \smallskip \\
 \quad \R{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\ 
 \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & \\ 
 \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ &  \\ 
 \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\ 
 \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & \smallskip   \\ 
 \hline
\end{tabular}
\caption{\hl{Blabla, mention that C is character, F is factor, I is integer, L is labelled, B is logical (boolean), N is numeric and D is Date. \newline Also: Check that everything in here is correct (i.e. corresponds to the output of allSummaryFunctions() etc}}
\label{table.SVCfunctions}
\end{table}

@@

@@

\hl{
\begin{itemize}
\item  
\end{itemize}
}

@@

Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable. In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described... blabla}. Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}. This section consists of three parts. We commence this section with an overview of how contents are controlled in \pkg{cleanR} in terms of what cleaning steps are performed. Secondly, we turn to the possibilities of customizing what the final cleanR file looks like. Lastly, we go through an example of how to put all these customization options together in practice.

\subsubsection{Writing a summaryFunction}
Though \pkg{cleanR} provides a special class for \code{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure. Specifically, they all follow the template below:
\begin{Verbatim}
mySummaryFunction <- function(v) {
	res <- [result of whatever summary we are doing]
	list(feature = "Feature name", result = res)
}
\end{Verbatim}
and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \code{summaryFunction} objects:
\begin{Verbatim}
mySummaryFunction <- summaryFunction(mySummaryFunction,
	description = "Some text describing what your summaryFunction does")
\end{Verbatim}
which adds the new function to the output of \code{allSummaryFunctions()}. Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one. In other words, e.g. integers are allowed, but matrices are not. Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).

\subsubsection{Writing a visualFunction}
\code{visualFunction}s are the functions that produce the figures of a \pkg{cleanR} output document. Writing a visual function is slightly more complicated than writing a summary function. This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \code{clean} to build standalone rmarkdown files. We recommend using the following structure:
\begin{Verbatim}
myVisualFunction <- function(v, vnam, doEval) {
	thisCall <- call("[the name of the function used to produce the plot]",
		v, [additional arguments to the plotting function])
	if (doEval) {
		return(eval(thisCall))
	} else return(deparse(thisCall)
}

myVisualFunction <- visualFunction(myVisualFunction,
	description = "Some text describing your visualFunction")
\end{Verbatim}
In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}). The latter \code{doEval} setting is not strictly necessary for its use in \code{clean}, but it makes it easier to assess what visualization options are available. In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them. As with \code{summaryFunction}s, an overview of all available \code{visualFunction}s in the environment can be obtained by calling
\begin{Verbatim}
allVisualFunctions()
\end{Verbatim}
and by calling
\begin{Verbatim}
allVisual(v, vnam, output = "html")
\end{Verbatim}
an overview of all plotting options applied on \code{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}

\subsubsection{Writing a checkFunction}
The last, but also most important, \pkg{cleanR} function type is the \code{checkFunction}. These are the functions that flag issues in the data and control the flow of the overall data cleaning process in the precheck stage. A \code{checkFunction} can be written using the following template:
\begin{Verbatim}
myCheckFunction <- function(v) {
	[do your check]
	problem <- [is there a problem? TRUE/FALSE]
	problemValues <- [vector of values in v that are problematic]
	problemStatus <- list(problem = problem, problemValues = problemValues)

	problemMessage <- "[The message that should be printed prior to listing
			problem values in the cleanR output]"

	outMessage <- messageGenerator(list(problem = problem,
		problemValues = problemValues, message = problemMessage))

	list(problem = problem, message = message) #problem is TRUE/FALSE,
		# message is a text string
}
myCheckFunction <- checkFunction(myCheckFunction,
	description = "[A description of your checkFunction]")
\end{Verbatim}
Only the input parameter (\code{v}) and the output format strictly has to follow this structure. However, we recommend using \code{messageGenerator} for consistent styling of all \code{checkFunction} messages. This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically. Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator} really is the easiest solution.

While the descriptions of \code{summaryFunction}s and \code{visualFunction}s are only for internal use in the \code{allSummaryFunctions()} and \code{allVisualFunctions()} outputs, respectively, \code{checkFunction} descriptions are actually visible in the \pkg{cleanR} output document. These are the brief descriptions presented in Part 1 \hl{eh? Is this clear? Whatever we call this section in the above} in the output document. If a \code{checkFunction} does not have a description (for instance, if it is just a regular function using the \code{checkFunction} input/output-structure), the function name will be printed instead of the description.

\begin{table}
\scriptsize
\begin{tabular}{p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}}
& \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
\hline
Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values  \\
Purpose & Describe some aspect of the variable, e.g. a central value, its dispersion or missingness. & Produce a distribution plot. & Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue. \\
Output (required) & A list with entries \newline \R{\$feature} Label for the summary value (as character string) \newline \R{\$result} The result of the summary (as character string) & A character string with \R{R} code for producing a plot. This code should be standalone, i.e. should include the data if necessary. & A list with entries \newline \R{\$problem} - a logical identifying whether an issue was found \newline \R{\$message} - character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
Output (recommended) & A \R{summaryResult} object (i.e. an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format). & \begin{tabular}{p{0.4\linewidth} | p{0.4\linewidth}} If \R{doEval} is \R{TRUE}: \newline A plot that will be opened by the graphic device in \R{R}. & If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above. \end{tabular} & A \R{checkResult} object (an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant. \\
Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
\end{tabular}
\caption{\hl{blabla. Fix formatting! Maybe use multirow and raggedRight stuff?}}
\label{table.functionTypes}
\end{table}

\hl{Eksempel}
\begin{Schunk}
\begin{Sinput}
> characterFoo <- function(v) {
+     if (substr(substitute(v), 1, 1) == "_") {
+         out <- list(problem=TRUE, message="Note that the variable name begins with \\_")
+     } else out <- list(problem=FALSE, message="")
+     out
+ }
> class(characterFoo) <- "checkFunction"
> attr(characterFoo, "description") <- "I really hate underscores"
> #clean(testData, characterChecks=c(defaultCharacterChecks(), "characterFoo"))
> 
\end{Sinput}
\end{Schunk}

\section{Something like examples}

@@

\subsection{Use cleanR for problem flagging}
If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be to not include all information about all variables. Or even for more reasonably sized datasets, sometimes a brief overview of the most pressing issues can be useful. This can be achieved by using the \R{onlyProblematic} argument in \R{clean()}. By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized. But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems, they might have. This can be done by also controlling the \R{mode} argument:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
\end{Sinput}
\end{Schunk}
Now only the checking results are printed, and only for variables where problems were identified. An even more minimal output can be generated by also leaving out the checking results - then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, onlyProblematic = TRUE, mode = NULL)
\end{Sinput}
\end{Schunk}
Of course, this can also be done without generating an overview document, by use of the \R{check()} function: 
\begin{Schunk}
\begin{Sinput}
> toyChecks <- check(toyData)
> foo <- function(x) {
>   any(sapply(x, function(y) y[["problem"]]))
> }
> sapply(toyChecks, foo)
\end{Sinput}
\begin{Soutput}
 var1  var2  var3  var4  var5  var6 
 TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
\end{Soutput}
\end{Schunk}
\hl{drop this last bit? too technical?}

\subsection{Include cleanR document in other files}

>>>>

\begin{figure}
\begin{center}
\frame{\includegraphics[width=7.5cm,page=2]{cleanR_toyData.pdf}} 
\frame{\includegraphics[width=7.5cm, page=3]{cleanR_toyData.pdf}}
%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
\end{center}
\label{fig:example1}
\caption{Example output from running \R{clean()} on the \R{toyData} dataset. First a summary of the full dataset,.. 
  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small.}}
\end{figure}

@@

By default, a pdf overview document is produced, saved to the disc (in the working directory) and opened for inspection. Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts. First, an overview of what was done is presented under the title \textit{Data cleaning summary}. Secondly, each variable in the dataset is presented in turn using (up to) three tools in the \textit{Variable list}: A table summarizing key features of the variable, a figure visualizing its distribution and potentially also a list of flagged issues. For instance, in the \R{numeric}-type variable \R{var2} from \R{toyData}, \R{clean()} has identified two values that are suspected to be miscoded missing values (\R{999} and \R{NaN}), while two values were also flagged as potential outliers that should be investigated more carefully. We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable. \hl{hm, delete the last part? or use different variable? Only these exact two checks were performed}. \\

\hl{metatext here: blabla more can be done blabla, in this section we present...}

\subsection{Controlling [something]}

\begin{table}
\small
\begin{tabular}{p{0.2\linewidth}p{0.5\linewidth}p{0.2\linewidth}}
\hline
Argument & Description & Default value \\
\hline

@@

\hline
\end{tabular}
\caption{A selection of commonly used arguments of \R{clean()}.  \hl{Note: I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions (dealt with in next subsection)}.  \hl{Fix formatting for \R{onlyProblematic}.}}
\label{table.cleanFormals}
\end{table}

\hl{Describe arguments in the table...}

@@

\begin{table}
\centering
\begin{tabular}{p{0.3\linewidth} p{0.35\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
 p{0.01\linewidth} p{0.01\linewidth}}
  \hline
& Description &  \multicolumn{7}{c}{Relevant variable classes} \\ \smallskip
 & &  C & F & I & L & B & N & D\\ 
  \hline \smallskip
  \textbf{\R{summaryFunction}s}  \smallskip \\
  \quad \R{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\ 
  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ &  \\ 
  \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \smallskip \\
 \textbf{\R{visualFunction}s} \smallskip \\
  \quad \R{basicVisual} & Histograms and barplots using graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \smallskip \\
 \textbf{\R{checkFunction}s} \smallskip \\
 \quad \R{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\ 
 \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & \\ 
 \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ &  \\ 
 \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\ 
 \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & \smallskip   \\ 
 \hline
\end{tabular}
\caption{\hl{Blabla, mention that C is character, F is factor, I is integer, L is labelled, B is logical (boolean), N is numeric and D is Date. \newline Also: Check that everything in here is correct (i.e. corresponds to the output of allSummaryFunctions() etc}}
\label{table.SVCfunctions}
\end{table}

@@

By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset. There are two stages in which this occurs \hl{(with ref. to figure/flowchart?)}:
\begin{enumerate}
\item In the precheck functions
\item In the summarize/visualize/check (SVC) step
\end{enumerate}
Each of these stages are controllable using appropriate function parameters in \code{clean}. In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs. However, if for instance the dataset at hand requires completely different visualizations, more control is needed. \pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s. They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes. By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands. Therefore, this section is dedicated to introducing each in turn.\\

@@

\hl{
\begin{itemize}
\item  Mention allCheckFunctions() etc. again here
\item Mention check(), visualize() and summarize() modes for data.frames
\end{itemize}
}

@@

\begin{table}[ht]
\small
\begin{tabular}{p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}}
& \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
\hline
Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values  \\
Purpose & Describe some aspect of the variable, e.g. a central value, its dispersion or missingness. & Produce a distribution plot. & Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue. \\
Output (required) & A list with entries \newline \R{\$feature} Label for the summary value (as character string) \newline \R{\$result} The result of the summary (as character string) & A character string with \R{R} code for producing a plot. This code should be standalone, i.e. should include the data if necessary. & A list with entries \newline \R{\$problem} - a logical identifying whether an issue was found \newline \R{\$message} - character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
Output (recommended) & A \R{summaryResult} object (i.e. an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format). & \begin{tabular}{p{0.4\linewidth} | p{0.4\linewidth}} If \R{doEval} is \R{TRUE}: \newline A plot that will be opened by the graphic device in \R{R}. & If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above. \end{tabular} & A \R{checkResult} object (an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant. \\
Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
\end{tabular}
\caption{\hl{blabla. Fix formatting! Maybe use multirow and raggedRight stuff?}}
\label{table.functionTypes}
\end{table}

Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable. In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions \hl{as described... blabla}. Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}. \hl{something about interactive mode. }

This section consists of three parts. First, we describe how the \R{clean} function can be extended by adding custom prechecks, summaries, checks and visualizations. In order to do this, one needs to produce small functions that obey to the syntax of \R{cleanR}, and therefore, the minimal requirements of this syntax are introduced. However, if one wishes to also use these custom functions in the interactive \R{cleanR} mode and e.g. have them appear in a \R{allSummaryFunctions()} call, have nice printing functions etc., one needs a few extra steps when producing the functions. These steps are presented in the second part of this section. Lastly, we turn to an example of how either type of custom cleaning functions can be produced, used in \R{clean()} and used interactively.

Throughout this section, Table~\ref{table.functionTypes} serves as an overview of the requirements and contents of each of the three function types used in \R{cleanR}, namely \R{summaryFunction}s (used for creating summaries), \R{visualFunction}s (used for creating plots) and \R{checkFunction}s (used for performing variable checks).

\subsection{Writing cleanR extensions: The minimal approach}

\subsection{Getting the full cleanR workhorse on your side}

\subsection{A worked example}

\section{Something like examples}
Finally, we will present a few examples of how to make \R{cleanR} solve certain issues related to data cleaning. First, we discuss the challenges related to cleaning large datasets, particularly in terms of memory use and computation speed. Next, we show how \R{cleanR} can be used for problem-flagging. Lastly, we discuss how the \R{cleanR} output document can be included in other \R{rmarkdown} documents as a mean to produce clear and concise documentation of a dataset.

@@

\subsection{Use cleanR for problem flagging}
If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be to not include all information about all variables. Or even for more reasonably sized datasets, sometimes a brief overview of the most pressing issues can be useful. This can be achieved by using the \R{onlyProblematic} argument in \R{clean()}. By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized. But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems, they might have. This can be done by also controlling the \R{mode} argument:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
\end{Sinput}
\end{Schunk}
Now only the checking results are printed, and only for variables where problems were identified. An even more minimal output can be generated by also leaving out the checking results - then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, onlyProblematic = TRUE, mode = NULL)
\end{Sinput}
\end{Schunk}
Of course, this can also be done without generating an overview document, by use of the \R{check()} function. When called on a \R{data.frame}, this function produces a list (of variables) of lists (of checks) of lists (or rather, \R{checkResult}s). Thus, the overall problem status of each variable can easily be unravelled using the list manipulation function \R{sapply()}: 
\begin{Schunk}
\begin{Sinput}
> toyChecks <- check(toyData)
> foo <- function(x) {
>   any(sapply(x, function(y) y[["problem"]]))
> }
> sapply(toyChecks, foo)
\end{Sinput}
\begin{Soutput}
 var1  var2  var3  var4  var5  var6 
 TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
\end{Soutput}
\end{Schunk}
and we find that only the final variable, \R{var6}, for which all observations have the value \R{"Irrelevant"}, is problem-free. \hl{drop this last bit? too technical?}

\subsection{Include cleanR document in other files}
Sometimes, a \R{cleanR} document might be a useful addition to a more general overview document, including also e.g. pairwise association plots, time series plots or exploratory analysis results. To this end, it is possible to produce a \R{cleanR} document that can readily be included in other \R{rmarkdown} files. This is done by using the \R{standAlone} argument in \R{clean}, which removes the preamble from the outputted \R{rmarkdown} file. Please note, that it is still necessary to indicate which \R{rmarkdown} type is being created; the pdf and html \R{rmarkdown} styles are unfortunately not identical.

If it is important that the embedded \R{cleanR} document can be rendered to either of these two file types, we recommend setting \R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby essentially removing almost all output type specific formatting code from the generated \R{rmarkdown} file.

On the other hand, if a pdf document is to be produced, the a few extra lines need to be added to the preamble of the master \R{rmarkdown} document - otherwise, the two-column layout code will produce an error. The following is an example of how such a master document preamble might look like and how the \R{cleanR\_toyData.Rmd} can be included:
\begin{Verbatim}
---
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---

```{r, child = 'cleanR_toyData.Rmd'}
```
\end{Verbatim}
\hl{Use proper formatting here.}

In the this example, the \R{cleanR\_toyData.Rmd} file could have been created as follows:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, standAlone = FALSE)
\end{Sinput}
\end{Schunk}
and the more minimal, html-style \R{rmarkdown} file described above can be produced using
\begin{Schunk}
\begin{Sinput}
> clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
\end{Sinput}
\end{Schunk}

\section{Conclusion/Concluding remarks/summary/?}
\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code....}
